Poker Chips, Earthquakes, and Computation

### 2014-10-11 finaldraft.tex
changed some notation stuff

In fact we are able to show that not only is this final board the same but that the number of earthquakes that target a particular square in order to stabilize the board is invariant of the order in which the earthquakes occur. We do so by first describing conditions under which the position of an Earthquake in a sequence can change without effecting the final state. We then define a process that rearranges a sequence of earthquakes without changing the final board state and produces the same sequence for any board.\\

@@

\begin{definition}
We say that an earthquake $E$ is valid if the square it targets has at least $4$ chips on it. 
We say that a sequence of earthquakes $E_1, E_2, \ldots E_n$ to be a valid sequence of earthquakes if each earthquake is valid in the order applied. That is 
\begin{equation*}
E^{m}[b, \{s_k\}](s_{m+1}  \geq 4
\end{equation*}
for all $1\geq m+1 \leq n$.
\end{definition}

@@

Now we can show that switching the order of two earthquakes with certain validity conditions does not affect the final board state.
\begin{lemma}
\label{swappinglemma}
Consider a board $b$ and two earthquakes $E_1$ and $E_2$, affecting squares $s_1$ and $s_2$ respectively. If $E_1, E_2$ is a valid sequence and $E_2$ is a valid earthquake when applied before $E_1$ then $E_2, E_1$ is also a valid sequence of earthquakes and 
\begin{equation*}
E_2[E_1[b, s_1] s_2] = E_1[LE_2[b, s_2], s_1] 
\end{equation*}
\end{lemma}
\begin{proof}
We will first show that $E_2, E_1$ is a valid sequence and then use this to show that the order in which the earthquakes occur does not matter.

We demonstrate this by contradiction. Assume  $E_2, E_1$ is an invalid sequence, that is $E_2[B, s_2](s_1)<4$. Since $E_1$ is a valid earthquake when applied directly to $b$ we know $b(s_1) \geq 4 $. 
Therefore we must have $b(s_1)<8$ and $s_2=s_1$. But then $E_1[b, s_1](s_1)<4$ and so $E_1, E_2$ would not be a valid sequence of earthquakes. This contradicts one our hypothesis and so $E_2, E_1$ is a valid sequence.

Now in order to demonstrate that the order in which the earthquakes are applied does not affect the final board state consider $E_2[ E_1[b, s_1], s_2](x)$ for some $(i,j) \in \mathbb{Z} \times \mathbb{Z}$. Applying Lemma ~\ref{earthquakeredefine} we have 
\begin{align*}
E_2[E_1[b, s_1], s_2](x) = E_1[b, s_1])(x)  + [[ x \in n(s_2) ]] - 4[[ x=s_2]] \\
= b(x) + [[ x \in n(s_1) ]] - 4[[x = s_1] + [[ x \in n(s_2) ]] - 4[[ x= s_2)]]  \\
= b(x) + [[ x \in n(s_2) ]] - 4[[x = s_2] + [[ x \in n(s_1) ]] - 4[[ x= s_1)]]\\
= E_2[B, s_2](x) + [[ x \in n(s_1) ]] - 4[[ x = s_1]] \\
= E_1[E_2 [b, s_2], s_1](x) 
\end{align*}
The critical step in this proof is recognizing that because $E_1$ and $E_2$ are valid earthquakes regardless of their ordering the algebra above was valid.
\end{proof}

We can use this lemma along with a simple induction to show that under appropriate conditions we can move an earthquake earlier in the sequence without affecting the final board state.
\begin{lemma}
Let $E_1,E_2,E_3,\dots, E_n$ be a sequence of earthquakes, where $E_k$ hits the square $s_k$. Consider  the sequence given by moving some earthquake $E_j$ to position $m<j$ in the sequence, such that $E_j$ is a valid earthquake for all positions between $m$ and $j$, inclusive. 
This sequence is valid and produces the same final board configuration.
\end{lemma}
\label{shiftlemma}
\begin{proof}
It is sufficient for us to show that the first $j$ earthquakes of each sequences are valid and that the board arrangement will be the same for both sequences after $j$ earthquakes. In fact we can also ignore the first $m-1$ terms, because their order is not changed so the board state will be identical after $m-1$ terms for both sequences. Let us call this board state $b_{m-1}$

>>>>

In fact we are able to show that not only is this final board the same but that the number of earthquakes that target a particular node in order to stabilize the board is invariant of the order in which the earthquakes occur. We do so by first describing conditions under which the position of an Earthquake in a sequence can change without effecting the final state. We then define a process that rearranges a sequence of earthquakes without changing the final board state and produces the same sequence for any board.\\

@@

\begin{definition}
We say that an earthquake $E$ is valid if the node it targets has at least $4$ chips on it. 
We say that a sequence of earthquakes $E_1, E_2, \ldots E_n$ to be a valid sequence of earthquakes if each earthquake is valid in the order applied. That is 
\begin{equation*}
E^{m}[b, \{s_k\}](s_{m+1}  \geq 4
\end{equation*}
for all $1\geq m+1 \leq n$.
\end{definition}

@@

Now we can show that switching the order of two earthquakes with certain validity conditions does not affect the final board state.
\begin{lemma}
\label{swappinglemma}
Consider a board $b$ and two earthquakes $E_1$ and $E_2$, affecting nodes $s_1$ and $s_2$ respectively. If $E_1, E_2$ is a valid sequence and $E_2$ is a valid earthquake when applied before $E_1$ then $E_2, E_1$ is also a valid sequence of earthquakes and 
\begin{equation*}
E_2[E_1[b, s_1] s_2] = E_1[LE_2[b, s_2], s_1] 
\end{equation*}
\end{lemma}
\begin{proof}
We will first show that $E_2, E_1$ is a valid sequence and then use this to show that the order in which the earthquakes occur does not matter.

We demonstrate this by contradiction. Assume  $E_2, E_1$ is an invalid sequence, that is $E_2[b, s_2](s_1)<4$. Since $E_1$ is a valid earthquake when applied directly to $b$ we know $b(s_1) \geq 4 $. 
Therefore we must have $b(s_1)<8$ and $s_2=s_1$. But then $E_1[b, s_1](s_1)<4$ and so $E_1, E_2$ would not be a valid sequence of earthquakes. This contradicts one our hypothesis and so $E_2, E_1$ is a valid sequence.

Now in order to demonstrate that the order in which the earthquakes are applied does not affect the final board state consider $E_2[ E_1[b, s_1], s_2](x)$ for some $x \in \mathbb{Z} \times \mathbb{Z}$. Applying Lemma ~\ref{earthquakeredefine} we have 
\begin{align*}
E_2[E_1[b, s_1], s_2](x) = E_1[b, s_1])(x)  + [[ x \in n(s_2) ]] - 4[[ x=s_2]] \\
= b(x) + [[ x \in n(s_1) ]] - 4[[x = s_1] + [[ x \in n(s_2) ]] - 4[[ x= s_2)]]  \\
= b(x) + [[ x \in n(s_2) ]] - 4[[x = s_2] + [[ x \in n(s_1) ]] - 4[[ x= s_1)]]\\
= E_2[b, s_2](x) + [[ x \in n(s_1) ]] - 4[[ x = s_1]] \\
= E_1[E_2 [b, s_2], s_1](x) 
\end{align*}
The critical step in this proof is recognizing that because $E_1$ and $E_2$ are valid earthquakes regardless of their ordering the algebra above was valid.
\end{proof}

We can use this lemma along with a simple induction to show that under appropriate conditions we can move an earthquake earlier in the sequence without affecting the final board state.
\begin{lemma}
Let $E_1,E_2,E_3,\dots, E_n$ be a sequence of earthquakes, where $E_k$ hits the node $s_k$. Consider  the sequence given by moving some earthquake $E_j$ to position $m<j$ in the sequence, such that $E_j$ is a valid earthquake for all positions between $m$ and $j$, inclusive. 
This sequence is valid and produces the same final board configuration.
\end{lemma}
\label{shiftlemma}
\begin{proof}
It is sufficient for us to show that the first $j$ earthquakes of each sequences are valid and that the board arrangement will be the same for both sequences after $j$ earthquakes. In fact we can also ignore the first $m-1$ terms, because their order is not changed so the board state will be identical after $m-1$ terms for both sequences. Let us call this board state $b_{m-1}$
